---
title: "Aspect 2 BERT Model — Condition Prediction"
author: "Wesley"
date: today
format: html
---

## Overview
This script builds the BERT-based model for Aspect 2 of the project. It uses the same prediction task as the benchmark model (aspect2_benchmark.qmd) — predicting the medical condition category from drug reviews — but replaces the Phase 2 linguistic features with 50 PCA-reduced BERT embeddings generated by Pete in Phase 2.5. The goal is to compare macro F1 against the benchmark to evaluate whether transfer learning improves condition prediction.
### Research Questions
This script addresses the same sub-research questions as the benchmark:
Sub-RQ4: Can the patient's medical condition be predicted from review text using BERT embeddings — and does it outperform simple linguistic features?
The BERT model directly answers whether semantic meaning captured by transfer learning adds predictive power beyond surface-level linguistic features. Results feed into Ellen's Phase 5 cross-aspect comparison.


## Model Design: Aspect 2 Benchmark vs Aspect 2 BERT
Both models tackle the same prediction task with the same train/test split and same cross-validation setup. The only difference is the input features:
Benchmark model (aspect2_benchmark.qmd):

Features: 6 linguistic features from Phase 2
Classifier: Multinomial logistic regression
Macro F1: 0.244

BERT model (aspect2_bert.qmd):

Features: 50 PCA components from BERT embeddings (pc_1 to pc_50)
Classifier: Random forest
Macro F1: to be determined

This is a deliberate controlled comparison — everything is kept identical except the features and classifier, isolating the effect of transfer learning on performance.


## Code chunk order

1. Setup & libraries
2. Load data
3. Join BERT embeddings (bert_pca_50.csv)
4. Filter Other & Unknown
5. Train/test split
6. Define features & trainControl
7. Train the model (random forest)
8. Load Model & Evaluate
9. Feature importance
10. Save outputs
11. PCA interpretation


## 1. Setup & Libraries

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(caret)
library(randomForest)
library(MLmetrics)

set.seed(42)
```


## 2. Load data

```{r}
# Load main dataset
df <- read_csv("data/drug_reviews_phase2.csv") %>%
  mutate(row_id = row_number())

# Load BERT PCA embeddings
emb_pca <- read_csv("data/bert_pca_50.csv")

# Load shared train indices
train_rows <- read_csv("data/shared_train_idx.csv")$row_id

# Quick check
cat("Main dataset rows:", nrow(df), "\n")
cat("BERT embeddings rows:", nrow(emb_pca), "\n")
cat("Train indices:", length(train_rows), "\n")
```


## 3. Join BERT Embeddings

```{r}
# Join BERT PCA embeddings to main dataset
df_full <- left_join(df, emb_pca, by = "row_id")

# Verify join
cat("Rows after join:", nrow(df_full), "\n")
cat("Columns after join:", ncol(df_full), "\n")

# Check pc columns are present
pc_cols <- paste0("pc_", 1:50)
cat("PCA columns present:", sum(pc_cols %in% names(df_full)), "/ 50\n")
```


## 4. Filter Other & Unknown

```{r}
# Filter out Other and Unknown categories
df_full <- df_full %>%
  filter(!condition_category %in% c("Other", "Unknown"))

# Convert condition_category to factor
df_full <- df_full %>%
  mutate(condition_category = as.factor(condition_category))

# Verify
cat("Rows after filtering:", nrow(df_full), "\n")
cat("Classes remaining:", n_distinct(df_full$condition_category), "\n")
```


## 5.  Train/Test Split

```{r}
# Split using Pete's shared train indices
df_train <- df_full %>% filter(row_id %in% train_rows)
df_test  <- df_full %>% filter(!row_id %in% train_rows)

# Clean factor levels
levels(df_train$condition_category) <- make.names(levels(df_train$condition_category))
levels(df_test$condition_category) <- make.names(levels(df_test$condition_category))

# Verify
cat("Train rows:", nrow(df_train), "\n")
cat("Test rows:", nrow(df_test), "\n")
cat("Classes in train:", n_distinct(df_train$condition_category), "\n")
cat("Classes in test:", n_distinct(df_test$condition_category), "\n")
```


## 6.  Define Features & trainControl

```{r}
# Define feature columns — pc_1 to pc_50
features <- paste0("pc_", 1:50)

# Prepare train and test sets
X_train <- df_train %>% select(all_of(features))
y_train <- df_train$condition_category

X_test  <- df_test %>% select(all_of(features))
y_test  <- df_test$condition_category

# Define cross-validation — identical to benchmark
ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = multiClassSummary,
  savePredictions = "final",
  allowParallel = TRUE
)
```


## 7. Train the Model

```{r}
#| eval: false

library(doParallel)
cl <- makePSOCKcluster(2)
registerDoParallel(cl)
```

```{r}
#| eval: false

# Train random forest model
model_bert <- train(
  x = X_train,
  y = y_train,
  method = "rf",
  trControl = ctrl,
  metric = "Accuracy",
  ntree = 500
)

# Save with different name to avoid overwrite
saveRDS(model_bert, "data/model_bert_500.rds")

print(model_bert)

stopCluster(cl)
```

### Note on Model Training Parameters
Initial training was attempted with ntree = 500 using parallel processing across multiple cores. However, this resulted in a memory allocation error (cannot allocate vector of size 1.2 Gb) on the Imperial College shared computing environment, likely due to concurrent usage by other users limiting available RAM. Parallel processing with doParallel requires each worker to hold its own copy of the data in memory, which combined with 50 PCA features across 155,120 training rows exceeded the available memory.
As a result, the final model was trained with ntree = 100 without parallel processing. This is a practical compromise — the performance difference between 100 and 500 trees is typically marginal (1-2% accuracy), and the core finding of BERT embeddings substantially outperforming the benchmark remains valid regardless of tree count.


## 8. Load Model & Evaluate

```{r}
# Load saved model
model_bert <- readRDS("data/model_bert.rds")

# Predict on test set
y_pred <- predict(model_bert, newdata = X_test)

# Confusion matrix
cm_bert <- confusionMatrix(y_pred, y_test)
print(cm_bert)

# Compute macro F1 manually
f1_per_class <- cm_bert$byClass[, "F1"]
macro_f1_bert <- mean(f1_per_class, na.rm = TRUE)

cat("\n--- BERT Model Results ---\n")
cat("Macro F1 (test):", round(macro_f1_bert, 4), "\n")
cat("Accuracy (test):", round(cm_bert$overall["Accuracy"], 4), "\n")
```


## 9. Feature Importance

```{r}
# Extract feature importance
importance_bert <- varImp(model_bert, scale = TRUE)
print(importance_bert)

# Plot feature importance — top 20 PCA components
plot(importance_bert,
     top = 20,
     main = "Feature Importance — BERT Model (Top 20 PCA Components)")
```


## 10. Save Outputs

```{r}
# Create plots folder if it doesn't exist
dir.create("plots", showWarnings = FALSE)

# Save feature importance plot
png("plots/aspect2_bert_importance.png", width = 1000, height = 800)
plot(importance_bert,
     top = 20,
     main = "Feature Importance — BERT Model (Top 20 PCA Components)")
dev.off()

# Save confusion matrix plot
png("plots/aspect2_bert_confusion.png", width = 1000, height = 800)
ggplot(as.data.frame(cm_bert$table),
       aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 3) +
  scale_fill_gradient(low = "white", high = "#4B9CD3") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Confusion Matrix — BERT Model",
       x = "Actual",
       y = "Predicted")
dev.off()

# Save metrics
sink("plots/aspect2_bert_metrics.txt")
cat("=== BERT Model Summary ===\n\n")
cat("Macro F1 (test):", round(macro_f1_bert, 4), "\n")
cat("Accuracy (test):", round(cm_bert$overall["Accuracy"], 4), "\n")
cat("Kappa (test):", round(cm_bert$overall["Kappa"], 4), "\n\n")
cat("Top 10 PCA components by importance:\n")
print(importance_bert$importance %>% arrange(desc(Overall)) %>% head(10))
sink()

cat("All outputs saved to /plots folder\n")
```


## 11. PCA Interpretation

```{r}
# --- First Try: Correlate top PCA components with known features ---
pca_correlations <- df_full %>%
  select(pc_10, pc_3, pc_7, sentiment_score, emotional_intensity,
         review_length, tfidf_specificity, rating) %>%
  cor() %>%
  round(3)

print(pca_correlations)

# --- Second Try: Read highest scoring reviews for pc_10 ---
cat("\n--- Top 5 reviews scoring highest on pc_10 ---\n")
df_full %>%
  arrange(desc(pc_10)) %>%
  select(review_clean, condition_category, pc_10) %>%
  head(5) %>%
  print()

cat("\n--- Top 5 reviews scoring lowest on pc_10 ---\n")
df_full %>%
  arrange(pc_10) %>%
  select(review_clean, condition_category, pc_10) %>%
  head(5) %>%
  print()
```

## Summary & Handoff Notes (for Ellen)

### What this file does
This script builds the BERT-based model for Aspect 2 of the project. It uses the same prediction task as the benchmark model — predicting medical condition category from drug reviews — but replaces the 6 Phase 2 linguistic features with 50 PCA-reduced BERT embeddings. The goal is to evaluate whether transfer learning improves condition prediction over simple linguistic features.

### Data

Input: drug_reviews_phase2.csv + bert_pca_50.csv — 193,895 rows after dropping Other and Unknown categories
Split: 80/20 train/test using shared_train_idx.csv — identical to benchmark for fair comparison
Target variable: condition_category — 14 clinical condition groups
Features used: pc_1 to pc_50 — 50 PCA components reduced from 768-dimensional BERT embeddings

### Model

Classifier: Random forest (method = "rf") via caret
Trees: ntree = 100 (500 attempted but failed due to memory constraints on shared computing environment)
Validation: 5-fold cross-validation
Optimisation metric: Accuracy during training, Macro F1 computed manually post-training

### Results

Macro F1 (test): 0.7931
Accuracy (test): 83.04%
Kappa: 0.7906

### Key Findings

BERT embeddings dramatically outperform Phase 2 linguistic features across all metrics — Macro F1 improved from 0.244 to 0.793, and accuracy from 32.65% to 83.04%
Unlike the benchmark where tfidf_specificity dominated at 100 with everything else below 1.5, BERT spreads importance across multiple components — pc_10 leads at 100 but pc_3 follows at 59.6 and pc_7 at 44.3
PCA components show near-zero correlation with all Phase 2 features (all below 0.25), confirming BERT captures fundamentally different linguistic information — deep semantic meaning rather than surface level features
Minority classes that were completely missed by the benchmark (Cancer, Cardiovascular, Neurological) are now predicted with reasonable sensitivity (0.64, 0.50, 0.68 respectively)

### Output Files
All files saved in the /plots folder:

aspect2_bert_metrics.txt — Macro F1, Accuracy, Kappa, top 10 PCA importance scores
aspect2_bert_confusion.png — 14x14 confusion matrix heatmap
aspect2_bert_importance.png — Top 20 PCA components importance chart

### Comparison with Benchmark for Phase 5
The BERT model substantially outperforms the benchmark across all metrics. Macro F1 improved from 0.244 to 0.793, an improvement of +0.549. Accuracy improved from 32.65% to 83.04%, an improvement of +50.39%. Kappa improved from 0.091 to 0.791, an improvement of +0.700.
Beyond raw numbers, the nature of feature importance also changed significantly. The benchmark relied almost entirely on a single feature — tfidf_specificity scoring 100 with everything else below 1.5. The BERT model distributes importance much more broadly — pc_10 leads at 100 but pc_3 follows at 59.6 and pc_7 at 44.3, suggesting BERT captures richer and more distributed linguistic signals.
This comparison is the key input for Ellen's Phase 5 cross-aspect analysis. The Macro F1 scores and feature importance rankings from both models should be compared against Evonne's Aspect 1 results to identify which linguistic signals are shared across both aspects and which are unique to clinical condition prediction.

### Limitations

ntree = 100 used instead of standard 500 due to memory constraints — results may improve slightly with more trees
PCA components are not directly interpretable — we cannot explicitly state what linguistic property pc_10 represents
Near-zero correlation between PCA components and Phase 2 features confirms BERT captures different information, but the exact nature of that information remains abstract
Random forest is less interpretable than multinomial logistic regression — tradeoff between performance and explainability