---
title: "Aspect 2 Side Effects Analysis"
author: "Wesley"
date: today
format: html
---

## Overview

This script addresses Sub-RQ2 of the project — detecting side effects from patient drug review text. Unlike the benchmark and BERT models which focus on condition prediction, this analysis takes a keyword-based approach to identify and quantify how frequently patients mention side effects in their reviews, and how this varies across condition categories and rating levels.
The approach is intentionally lightweight — rather than building a full classifier, we define a curated list of 30-50 common side effect terms, flag reviews and sentences containing them, and analyse the resulting frequency patterns. This is sufficient to answer Sub-RQ2 and provides clinically meaningful insights without requiring additional model training.

### Research Questions

Sub-RQ2: Can side effects be detected from patient review text using linguistic patterns?
This script answers Sub-RQ2 by showing how frequently side effect language appears across different condition categories and rating levels. The key questions are:

Which condition categories mention side effects most frequently?
Do lower rated reviews mention more side effects than higher rated ones?
Is there a relationship between side effect mentions and patient satisfaction?

The findings feed directly into Ellen's Phase 5 interpretation — specifically the clinical implications section around auto-detecting side effects from review text.


## 1. Setup & libraries

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(tidytext)
library(ggplot2)

set.seed(42)
```


## 2. Load data

```{r}
# Load main dataset
df <- read_csv("data/drug_reviews_phase2.csv") %>%
  mutate(row_id = row_number())

# Filter Other and Unknown — consistent with Tasks A and B
df <- df %>%
  filter(!condition_category %in% c("Other", "Unknown"))

# Verify
cat("Total rows:", nrow(df), "\n")
cat("Condition categories:", n_distinct(df$condition_category), "\n")
```


## 3. Define side effect keywords

```{r}
# Curated list of common side effect terms
side_effects <- c(
  # Gastrointestinal
  "nausea", "vomiting", "diarrhea", "constipation", "stomach",
  "bloating", "indigestion", "heartburn", "cramps", "appetite",
  
  # Neurological
  "headache", "dizziness", "drowsiness", "fatigue", "insomnia",
  "tremor", "confusion", "memory", "concentration", "migraine",
  
  # Psychological
  "anxiety", "depression", "mood", "irritability", "suicidal",
  "hallucination", "paranoia", "agitation", "panic",
  
  # Physical
  "rash", "itching", "swelling", "weight", "sweating",
  "bleeding", "bruising", "pain", "weakness", "fever",
  
  # Cardiovascular
  "palpitation", "heartbeat", "blood pressure", "chest",
  
  # Sexual
  "libido", "sexual", "erectile"
)

cat("Total side effect terms:", length(side_effects), "\n")
```


## 4. Flag side effect sentences

```{r}
# Create a regex pattern from side effect terms
pattern <- paste(side_effects, collapse = "|")

# Flag reviews containing any side effect term
df <- df %>%
  mutate(
    has_side_effect = str_detect(tolower(review_clean), pattern),
    side_effect_count = str_count(tolower(review_clean), pattern)
  )

# Quick summary
cat("Reviews mentioning side effects:", sum(df$has_side_effect), "\n")
cat("Percentage:", round(mean(df$has_side_effect) * 100, 1), "%\n")
cat("Average mentions per review:", round(mean(df$side_effect_count), 2), "\n")
```
### Findings: Side Effect Detection
The keyword matching approach flagged 138,907 reviews (71.6% of all reviews) as containing at least one side effect term, with an average of 1.9 mentions per review. This directly answers Sub-RQ2 — side effect language is pervasive in patient drug reviews and can be reliably detected using keyword matching.
However, a key limitation of this approach is the potential for false positives. Simple pattern matching may match partial words — for example "pain" matching "Spain", or "mood" matching "childhood". A proper NLP classifier would reduce these errors but is beyond the scope of this analysis. The 71.6% figure should therefore be interpreted as an upper bound estimate rather than a precise measurement.


## 5. Frequency by condition

```{r}
# Side effect frequency by condition category
condition_side_effects <- df %>%
  group_by(condition_category) %>%
  summarise(
    total_reviews = n(),
    reviews_with_se = sum(has_side_effect),
    pct_with_se = round(mean(has_side_effect) * 100, 1),
    avg_se_count = round(mean(side_effect_count), 2)
  ) %>%
  arrange(desc(pct_with_se))

print(condition_side_effects)

# Visualise
plot_condition <- ggplot(condition_side_effects, 
       aes(x = reorder(condition_category, pct_with_se), 
           y = pct_with_se,
           fill = pct_with_se)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient(low = "#99CC99", high = "#CC6666") +
  theme_minimal() +
  labs(title = "Side Effect Mention Rate by Condition Category",
       x = "Condition Category",
       y = "% Reviews Mentioning Side Effects",
       fill = "Rate %")

print(plot_condition)
```


## 6. Frequency by rating

```{r}
# Side effect frequency by rating level
rating_side_effects <- df %>%
  group_by(rating) %>%
  summarise(
    total_reviews = n(),
    reviews_with_se = sum(has_side_effect),
    pct_with_se = round(mean(has_side_effect) * 100, 1),
    avg_se_count = round(mean(side_effect_count), 2)
  ) %>%
  arrange(rating)

print(rating_side_effects)

# Visualise
plot_rating <- ggplot(rating_side_effects,
       aes(x = rating, y = pct_with_se, fill = pct_with_se)) +
  geom_col() +
  scale_x_continuous(breaks = 1:10) +
  scale_fill_gradient(low = "#99CC99", high = "#CC6666") +
  theme_minimal() +
  labs(title = "Side Effect Mention Rate by Rating Level",
       x = "Patient Rating (1-10)",
       y = "% Reviews Mentioning Side Effects",
       fill = "Rate %")

print(plot_rating)
```


## 7. Save outputs

```{r}
dir.create("plots", showWarnings = FALSE)

# Save condition plot
ggsave("plots/aspect2_se_by_condition.png", plot = plot_condition, width = 10, height = 8)

# Save rating plot
ggsave("plots/aspect2_se_by_rating.png", plot = plot_rating, width = 10, height = 6)

# Save summary stats
sink("plots/aspect2_se_metrics.txt")
cat("=== Side Effects Analysis Summary ===\n\n")
cat("Total reviews analysed:", nrow(df), "\n")
cat("Reviews with side effects:", sum(df$has_side_effect), "\n")
cat("Overall percentage:", round(mean(df$has_side_effect) * 100, 1), "%\n\n")
cat("By condition category:\n")
print(condition_side_effects)
cat("\nBy rating:\n")
print(rating_side_effects)
sink()

cat("All outputs saved to /plots folder\n")
```


## Summary & Handoff Notes (for Ellen)

### What this file does
This script addresses Sub-RQ2 of the project — detecting side effects from patient drug review text. Using a curated list of 46 common side effect terms, it flags reviews containing side effect language and analyses frequency patterns across condition categories and rating levels. This is a keyword-based approach rather than a full classifier, which is sufficient to answer Sub-RQ2 and provide clinically meaningful insights.

### Data

Input: drug_reviews_phase2.csv — 193,895 rows after dropping Other and Unknown categories
No train/test split — full dataset used for frequency analysis
No embeddings needed — keyword matching only

### Method

A list of 46 side effect terms was curated across six clinical categories — gastrointestinal, neurological, psychological, physical, cardiovascular, and sexual. Reviews were flagged using str_detect on the cleaned review text, with str_count used to count total mentions per review. All matching was case insensitive.

### Key Findings

#### Overall:

138,907 reviews (71.6%) mention at least one side effect term
Average of 1.9 side effect mentions per review
Side effect language is pervasive across all condition categories

#### By condition category:

Pain & Inflammation has the highest side effect mention rate at 85.9%, likely reflecting the nature of chronic pain treatments and their systemic effects
Respiratory has the lowest rate at 38.0%, suggesting patients focus more on symptom relief than side effects
Most conditions cluster between 60-80%, indicating side effect language is consistently common regardless of condition

#### By rating level:

Higher rated reviews (9-10) mention fewer side effects — 71.6% and 64.5% respectively
Peak side effect mentions occur at rating 4 (80.1%), not rating 1 as might be expected
This suggests patients who are satisfied with their treatment mention side effects less, while moderately dissatisfied patients are most likely to document side effects in detail

### Output Files

All files saved in the /plots folder:

aspect2_se_metrics.txt — full frequency tables by condition and rating
aspect2_se_by_condition.png — side effect rate bar chart by condition category
aspect2_se_by_rating.png — side effect rate bar chart by rating level

### Connection to Other Aspects

The finding that higher ratings correlate with fewer side effect mentions directly connects to Evonne's Aspect 1 work on review usefulness and sentiment. Ellen should examine whether reviews flagged as high side effect mentions are also rated as more useful — this would suggest that clinical detail increases perceived usefulness, bridging the patient and clinical perspectives.

### Limitations

Keyword matching may produce false positives — common words like "pain", "mood", and "chest" may match non-side-effect contexts
The 71.6% figure should be treated as an upper bound estimate
No sentence-level analysis was performed — the entire review was treated as a single unit
Keyword list was manually curated and may miss domain-specific terminology used by patients

