---
title: "Drug Review Dataset — Phase 0 and Phase 1: EDA"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)
```

# Overview

This notebook is to describe the **Exploratory Data Analysis**. It loads the raw train/test CSVs, performs minimal cleaning, then explores the dataset through two research aspects:

-   **Data Loading & Cleaning**: Load raw CSVs, remove duplicates and invalid rows, parse dates
-   **EDA - General Data Overview**: Dataset-level summaries (rating, review length, useful votes, conditions)
-   **EDA Part A - Patient Perspective**: *"What makes a drug review persuasive and useful to other patients?"* Relationships between review attributes and usefulness, ending with a LASSO word-coefficient model.
-   **EDA Part B - Clinical Perspective**: *"What can we extract from reviews to support medical understanding?"* Side-effect mining with a condition-exclusion rule, everyday language vs clinical language analysis, and condition vocabulary exploration.

# Setup: Install & Load Packages

```{r packages}
required_packages <- c(
  "tidyverse",
  "tidytext",
  "textclean",
  "lubridate",
  "quanteda",
  "quanteda.textstats",
  "quanteda.textplots",
  "ggplot2",
  "scales",
  "glmnet",
  "ggrepel"
)

installed <- rownames(installed.packages())
for (pkg in required_packages) {
  if (!pkg %in% installed) install.packages(pkg, dependencies = TRUE)
}

library(tidyverse)
library(tidytext)
library(textclean)
library(lubridate)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(ggplot2)
library(scales)
library(glmnet)
library(ggrepel)

source("TAB_dfm.R")
source("kendall_acc.R")
```

# Phase 0: Data Loading & Cleaning

## Step 1: Load & Combine Train and Test

```{r load-combine}
train_raw <- read.csv("drugsComTrain_raw.csv", stringsAsFactors = FALSE)
test_raw  <- read.csv("drugsComTest_raw.csv",  stringsAsFactors = FALSE)

train_raw$split <- "train"
test_raw$split  <- "test"

df <- rbind(train_raw, test_raw)
cat("Combined rows:", nrow(df), "\n")
cat("Columns:", paste(names(df), collapse = ", "), "\n")
head(df, 5)
```

## Step 2: Remove Duplicates

```{r remove-duplicates}
n_before <- nrow(df)
df <- df %>%
  distinct(uniqueID, .keep_all = TRUE) %>%
  distinct(drugName, review, .keep_all = TRUE)

cat("Rows removed:", n_before - nrow(df), "\n")
cat("Rows remaining:", nrow(df), "\n")
```

## Step 3: Strip HTML from Reviews

```{r strip-html}
df$review <- df$review %>%
  gsub('&#039;', "'", .) %>%
  gsub('&amp;',  '&', .) %>%
  gsub('&quot;', '"', .) %>%
  gsub('&lt;',   '<', .) %>%
  gsub('&gt;',   '>', .) %>%
  gsub('&nbsp;', ' ', .) %>%
  gsub('<[^>]+>', ' ', .) %>%
  str_squish()
```

## Step 4: Lowercase

```{r lowercase}
df$review_clean <- tolower(df$review)
```

## Step 5: Filter Short Reviews (\< 20 words)

```{r filter-short}
df$word_count_raw <- str_count(df$review_clean, "\\S+")
n_before <- nrow(df)
df <- df %>% filter(word_count_raw >= 20)

cat("Rows removed (too short):", n_before - nrow(df), "\n")
cat("Rows remaining:", nrow(df), "\n")
```

## Step 6: Handle Missing Conditions

```{r missing-conditions}
df$condition[is.na(df$condition) | df$condition == ""] <- "Unknown"
```

## Step 7: Parse Date Column

```{r parse-date}
df$date <- dmy(df$date)
cat("Date range:", format(min(df$date)), "to", format(max(df$date)), "\n")
```

## Step 8: Validate Rating Range

```{r validate-rating}
df <- df %>% filter(rating >= 1 & rating <= 10)
cat("Cleaned dataset:", nrow(df), "rows x", ncol(df), "columns\n")
```

------------------------------------------------------------------------

# EDA: General Data Overview

A standard look at the shape of the cleaned dataset.

## Rating Distribution

```{r rating-dist}
ggplot(df, aes(x = factor(rating))) +
  geom_bar(fill = "#2c7bb6", colour = "white") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Rating Distribution",
    x     = "Star Rating (1-10)",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

Heavy right-skew, ratings of 9 and 10 dominate.

## Review Length

```{r review-length-dist}
df <- df %>%
  mutate(n_words = str_count(review_clean, "\\S+"))

ggplot(df, aes(x = n_words)) +
  geom_histogram(binwidth = 10, fill = "#66c2a5", colour = "white") +
  coord_cartesian(xlim = c(0, 300)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Distribution of Review Length",
    x     = "Number of Words",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

Most reviews fall between 25 and 150 words, with a clear peak around 125–150 words that likely reflects the platform's review form size. Very short reviews were already filtered out at the cleaning stage, and reviews above 200 words are rare — suggesting patients tend to write compact but structured accounts rather than long-form narratives.

## Useful Votes

```{r useful-dist}
ggplot(df, aes(x = usefulCount)) +
  geom_histogram(binwidth = 5, fill = "#fc8d62", colour = "white") +
  coord_cartesian(xlim = c(0, 150)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Distribution of usefulCount",
    x     = "Useful Vote Count",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

The plot shows a heavy right-skew where the vast majority of reviews receive fewer than 25 useful votes, while a long tail of reviews accumulate 50, 100, or more. This is typical of engagement data in which a small number of reviews attract disproportionate attention.

## Reviews Over Time

```{r time-dist}
df %>%
  mutate(year = year(date)) %>%
  count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = "#8da0cb", colour = "white") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2008, 2017, by = 1)) +
  labs(
    title = "Number of Reviews by Year",
    x     = "Year",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

Review volume grows steadily from 2008 through 2014, then jumps dramatically in 2015–2017. This is important context: the dataset is weighted toward recent reviews, so any model we build reflects more recent patient vocabulary and platform norms than older ones.

## Top Conditions

```{r top-conditions}
top_conditions <- df %>%
  count(condition, sort = TRUE) %>%
  slice_head(n = 15)

ggplot(top_conditions, aes(x = reorder(condition, n), y = n)) +
  geom_col(fill = "#5e81ac") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 15 Conditions by Number of Reviews",
    x     = NULL,
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

Birth Control dominates with over 35,000 reviews, which is roughly three times more than the second-place Depression. Mental health conditions (Depression, Anxiety, Bipolar Disorder, ADHD, Insomnia) make up a large share of the rest, alongside chronic conditions like Pain, Weight Loss/Obesity, and Diabetes.

## Top Drugs

```{r top-drugs}
top_drugs_overall <- df %>%
  count(drugName, sort = TRUE) %>%
  slice_head(n = 15)

ggplot(top_drugs_overall, aes(x = reorder(drugName, n), y = n)) +
  geom_col(fill = "#b5838d") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 15 Drugs by Number of Reviews",
    x     = NULL,
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

------------------------------------------------------------------------

# EDA Part A: Patient Perspective {.tabset}

> **Research question:** *"What makes a drug review persuasive and useful to other patients?"*
>
> We explore how basic review attributes relate to `usefulCount`, then turn to the review text itself with a LASSO word-coefficient model.

## A1. Review Length vs Usefulness

First, we want to explore if longer, more detailed reviews earn more useful votes.

```{r a1-length-useful}
df <- df %>%
  mutate(
    length_group = cut(
      n_words,
      breaks = c(0, 50, 100, 150, 200, Inf),
      labels = c("20-50", "51-100", "101-150", "151-200", "200+"),
      include.lowest = TRUE
    )
  )

ggplot(df, aes(x = length_group, y = usefulCount)) +
  geom_boxplot(fill = "skyblue", outlier.alpha = 0.08) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    title = "Useful Votes by Review Length",
    x     = "Review Length (words)",
    y     = "usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

Median useful votes stay roughly flat across the 20-200 word range (around 15-18 votes), but the 200+ group jumps sharply with median around 26 votes with a much wider spread. The 151-200 group is actually slightly *lower* than shorter reviews, suggesting it's not length alone that matters. Short punchy reviews and very long detailed reviews both do well, while medium-long reviews add words without adding signal.

## A2. Rating vs Usefulness

We evaluate if reviews with extreme ratings attract more votes than moderate ones.

```{r a2-rating-useful}
ggplot(df, aes(x = factor(rating), y = usefulCount)) +
  geom_boxplot(fill = "#fdb462", outlier.alpha = 0.08) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    title = "Useful Votes by Star Rating",
    x     = "Star Rating",
    y     = "usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

```{r a2-rating-useful-summary}
df %>%
  group_by(rating) %>%
  summarise(
    n             = n(),
    avg_useful    = round(mean(usefulCount, na.rm = TRUE), 1),
    median_useful = median(usefulCount, na.rm = TRUE),
    .groups       = "drop"
  )
```

The pattern is monotonic, where average useful votes climb steadily from 15.5 at rating 1 to 38.4 at rating 10. Rather than extreme ratings attracting the most engagement, *positive* reviews earn substantially more useful votes than negative ones. It's also worth noting this could partly reflect selection, where highly-rated drugs are prescribed more often, so more users read their reviews and vote.

## A3. Usefulness Over Time

Older reviews have had longer to accumulate votes. This is a purely descriptive check to interpret `usefulCount`.

```{r a3-time-useful}
df %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarise(
    n          = n(),
    avg_useful = mean(usefulCount, na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  ggplot(aes(x = year, y = avg_useful)) +
  geom_col(fill = "#e76f51") +
  scale_x_continuous(breaks = seq(2008, 2017, by = 1)) +
  labs(
    title = "Average usefulCount by Review Year",
    subtitle = "Older reviews have had longer to accumulate votes",
    x     = "Year",
    y     = "Average usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

A clear exposure-bias pattern is seen. Reviews from 2008 average \~66 useful votes, but reviews from 2017 average only \~12. This does not mean 2008 reviews were better written but they've simply had nearly a decade to accumulate votes, while 2017 reviews had weeks. This matters for interpreting the LASSO model below, which any word that was more common in older reviews will get an inflated coefficient because those reviews have higher vote counts by virtue of age. We acknowledge this as a limitation during the EDA.

## A4. LASSO Word Coefficients: Which Words Predict Usefulness?

A LASSO regression on the document-feature matrix identifies which individual words are *predictive* of `usefulCount` after accounting for all other terms simultaneously. Words shrunk to zero are uninformative; non-zero positive coefficients reveal the vocabulary of usefulness. We use a 30k sample with a train/test split for accuracy evaluation.

```{r a4-lasso-train}
set.seed(42)
lasso_sample <- df %>% sample_n(min(nrow(df), 30000))

train_idx   <- sample(1:nrow(lasso_sample), 0.8 * nrow(lasso_sample))
lasso_train <- lasso_sample[train_idx, ]
lasso_test  <- lasso_sample[-train_idx, ]

dfm_train <- TAB_dfm(lasso_train$review_clean, ngrams = 1)
dfm_test  <- TAB_dfm(lasso_test$review_clean, ngrams = 1, min.prop = 0) %>%
  dfm_match(colnames(dfm_train))

rev_model <- cv.glmnet(
  x = as.matrix(dfm_train),
  y = lasso_train$usefulCount
)
```

```{r a4-lasso-plot, fig.height=8, fig.width=11}
plotDat <- rev_model %>%
  coef() %>%
  drop() %>%
  as.data.frame() %>%
  rownames_to_column(var = "ngram") %>%
  rename(score = ".") %>%
  filter(score != 0 & ngram != "(Intercept)" & !is.na(score)) %>%
  left_join(
    data.frame(
      ngram = colnames(dfm_train),
      freq  = colMeans(dfm_train)
    ),
    by = "ngram"
  )

cat("Non-zero LASSO coefficients:", nrow(plotDat),
    "(", sum(plotDat$score > 0), "positive,",
        sum(plotDat$score < 0), "negative )\n")

plotDat %>%
  ggplot(aes(x = score, y = freq, label = ngram)) +
  geom_vline(xintercept = 0) +
  # Negative coefficients: grey points, no labels
  geom_point(
    data   = plotDat %>% filter(score < 0),
    colour = "grey70",
    size   = 3,
    alpha  = 0.7
  ) +
  # Positive coefficients: green gradient
  geom_point(
    data = plotDat %>% filter(score >= 2),
    aes(colour = score),
    size  = 3,
    alpha = 0.8
  ) +
  geom_label_repel(
    data         = plotDat %>% filter(score >= 2),
    aes(colour   = score),
    fill         = "white",
    size         = 5,
    max.overlaps = 30,
    force        = 6
  ) +
  scale_colour_gradient(low = "#a8dadc", high = "#1b4332") +
  scale_y_continuous(trans  = "log2",
                     breaks = c(.01, .05, .1, .2, .5, 1, 2, 5)) +
  theme_bw() +
  labs(
    title = "LASSO Word Coefficients — Predicting usefulCount",
    x     = "Coefficient in Model",
    y     = "Uses per Review"
  ) +
  theme(legend.position = "none",
        plot.title      = element_text(size = 18, face = "bold"),
        axis.title      = element_text(size = 20),
        axis.text       = element_text(size = 16))
```

```{r a4-lasso-accuracy}
test_pred <- predict(rev_model, newx = as.matrix(dfm_test))[, 1]
acc_ngram <- kendall_acc(lasso_test$usefulCount, test_pred)
acc_ngram
```

The LASSO shrinks most words to zero and keeps only those that carry signal. We have 132 non-zero coefficients (73 positive, 59 negative) out of thousands of candidate terms. A few patterns stand out:

-   **Specific drug names and conditions drive usefulness.** The strongest positive coefficients come from named medications (*wellbutrin*, *cymbalta*, *prozac*, *zoloft*) and specific conditions (*fibromyalgia*, *panic*, *disord* for "disorder"). Reviews that name exactly what they're treating are more useful to other patients.
-   **Outcome language matters.** Words like *miracl* (miracle), *life*, *calm*, *energi* (energy), *save*, and *famili* (family) all have strong positive coefficients. These signal reviews that describe the *impact* a drug had on daily life, not just whether it worked.
-   **Emotional specificity beats generic praise.** *Amaz* (amazing), *happi*, and *panic* sit far right, while generic *great* is much closer to zero. This suggests that descriptive experience earns more votes.
-   **"Without" as a positive signal** is interesting. It likely appears in phrases like "without this drug, I..." or "without side effects," both of which frame concrete before/after stories.

Kendall's tau on the held-out test set is \~67%, meaning the model correctly ranks two random reviews by usefulness about two-thirds of the time.

**Caveat on `usefulCount` as outcome:** As shown in A3, `usefulCount` is cumulative and biased toward older reviews. Some of these coefficients may partly reflect vocabulary common in earlier years rather than pure usefulness signal. Medical vocabulary is reasonably stable over 2008–2017, so this likely affects absolute coefficient values more than the ranking of top words.

------------------------------------------------------------------------

# EDA Part B: Clinical Perspective {.tabset}

We want to address the following question:*"What can we extract from reviews to support medical understanding?"*

## B1. Side-Effect Detection with Condition-Exclusion

A symptom term like "depression" could be the patient's *condition* or a *side effect* of their medication. We address this with a **condition-exclusion rule**, such that a term only counts as a side effect if it doesn't match the patient's listed condition.

```{r b1-side-effect-detection}
# Categorised side-effect dictionary
side_effect_dict <- tibble(
  category = c(
    rep("Gastrointestinal", 4),
    rep("Neurological",     4),
    rep("Psychiatric",      4),
    rep("Metabolic",        3),
    rep("Dermatological",   3),
    rep("Musculoskeletal",  2)
  ),
  term = c(
    "nausea", "vomiting", "diarrhea", "constipation",
    "headache", "dizziness", "drowsiness", "tremor",
    "anxiety", "depression", "insomnia", "mood swings",
    "weight gain", "weight loss", "appetite",
    "rash", "itching", "acne",
    "cramps", "pain"
  )
)

condition_lower <- tolower(df$condition)

# Vectorised detection matrices
term_in_review <- sapply(side_effect_dict$term, function(t) {
  str_detect(df$review_clean, regex(t, ignore_case = TRUE))
})
term_in_condition <- sapply(side_effect_dict$term, function(t) {
  str_detect(condition_lower, regex(t, ignore_case = TRUE))
})
side_effect_matrix <- term_in_review & !term_in_condition

df$n_side_effects   <- rowSums(side_effect_matrix)
df$has_side_effects <- df$n_side_effects > 0
```

```{r b1-side-effect-overview}
ggplot(df, aes(x = has_side_effects)) +
  geom_bar(fill = "#6a994e") +
  scale_x_discrete(labels = c("FALSE" = "No side effects detected",
                               "TRUE"  = "Side effects detected")) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Reviews with Detected Side Effects (condition-excluded)",
    x     = NULL,
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)

df %>%
  group_by(has_side_effects) %>%
  summarise(
    n          = n(),
    pct        = round(n() / nrow(df) * 100, 1),
    avg_rating = round(mean(rating, na.rm = TRUE), 2),
    .groups    = "drop"
  )
```

Approximately 52% of reviews mention at least one side-effect term after condition-exclusion. The rating comparison shows reviews that mention side effects average 6.64 stars, while those that don't average 7.35. That approximately 0.7-star gap is not huge, but confirms that side-effect reporting is genuinely tied to worse experiences rather than being random noise. If we see this from another perspective, if we see side-effect language in a review, we can expect a slightly lower star rating. This validates the detection rule as capturing real clinical signal rather than just everyday vocabulary.

## B2. Side-Effect Profile by Drug (Heatmap)

For the top 10 most-reviewed drugs, we explore which side effects appear most frequently. The condition-exclusion rule ensures "pain" in a Pain review doesn't inflate the count for pain medications.

```{r b2-side-effect-heatmap, fig.height=7}
top_drugs <- df %>%
  count(drugName, sort = TRUE) %>%
  slice_head(n = 10) %>%
  pull(drugName)

drug_se_data <- df %>%
  filter(drugName %in% top_drugs) %>%
  select(drugName, condition, review_clean)

drug_term_detected <- sapply(side_effect_dict$term, function(t) {
  str_detect(drug_se_data$review_clean, regex(t, ignore_case = TRUE))
})
drug_cond_match <- sapply(side_effect_dict$term, function(t) {
  str_detect(tolower(drug_se_data$condition), regex(t, ignore_case = TRUE))
})
drug_se_valid <- drug_term_detected & !drug_cond_match

heatmap_grid <- expand.grid(
  drug = top_drugs,
  term = side_effect_dict$term,
  stringsAsFactors = FALSE
) %>%
  left_join(side_effect_dict, by = "term")

heatmap_grid$pct <- map2_dbl(heatmap_grid$drug, heatmap_grid$term, function(d, t) {
  idx <- which(drug_se_data$drugName == d)
  col <- which(side_effect_dict$term == t)
  round(sum(drug_se_valid[idx, col]) / length(idx) * 100, 1)
})

ggplot(heatmap_grid, aes(x = term, y = drug, fill = pct)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = paste0(pct, "%"),
                colour = pct > 25),
            size = 2.8) +
  scale_fill_gradient(low = "#f1faee", high = "#1b4332", name = "% of reviews") +
  scale_colour_manual(values = c("FALSE" = "black", "TRUE" = "white"),
                      guide  = "none") +
  labs(
    title = "Side-Effect Mentions by Drug",
    x     = NULL,
    y     = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This demonstrates that it is possible to mine clinical signal. Without any labelled training data, the method produces drug-specific side-effect summaries that a prescribing clinician could actually use.

## B3. Everyday Language vs Clinical Language Gap

Patients describe symptoms in everyday language; doctors use clinical terms. We quantify this gap to motivate why NLP is needed to bridge both registers.

```{r b3-language-gap}
language_map <- tibble(
  informal = c("can't sleep", "feeling down", "throwing up", "brain fog",
               "tummy ache", "feeling sick", "gained weight", "lost weight",
               "breaking out", "feeling anxious", "dizzy"),
  clinical = c("insomnia", "depression", "vomiting", "cognitive impairment",
               "abdominal pain", "nausea", "weight gain", "weight loss",
               "acne", "anxiety", "dizziness")
)

language_map$informal_count <- sapply(language_map$informal, function(t) {
  sum(str_detect(df$review_clean, regex(t, ignore_case = TRUE)))
})
language_map$clinical_count <- sapply(language_map$clinical, function(t) {
  sum(str_detect(df$review_clean, regex(t, ignore_case = TRUE)))
})

language_long <- language_map %>%
  pivot_longer(cols = c(informal_count, clinical_count),
               names_to = "type", values_to = "count") %>%
  mutate(
    type  = if_else(type == "informal_count", "Informal", "Clinical"),
    label = if_else(type == "Informal", informal, clinical)
  )

ggplot(language_long, aes(x = reorder(informal, -count), y = count, fill = type)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Informal" = "#e9c46a", "Clinical" = "#264653")) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Layperson vs Clinical Term Frequency",
    x     = "Symptom (informal label)",
    y     = "Number of Reviews",
    fill  = "Language Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Clinical terms dominate most pairs. This suggests patients on Drugs.com are relatively medically literate and mostly use clinical vocabulary.

## B4. Word Clouds: High vs Low Rated Reviews

```{r b4-wordclouds, fig.height=6, fig.width=12}
df <- df %>%
  mutate(
    rating_group = case_when(
      rating <= 4 ~ "Low",
      rating <= 7 ~ "Medium",
      rating >= 8 ~ "High"
    )
  )

dfm_high <- TAB_dfm(
  df %>% filter(rating_group == "High") %>% pull(review_clean),
  ngrams     = 1:2,
  stop.words = TRUE,
  min.prop   = 0.01
)

dfm_low <- TAB_dfm(
  df %>% filter(rating_group == "Low") %>% pull(review_clean),
  ngrams     = 1:2,
  stop.words = TRUE,
  min.prop   = 0.01
)

par(mfrow = c(1, 2))
textplot_wordcloud(dfm_high, max_words = 75)
title("High-rated Reviews")
textplot_wordcloud(dfm_low, max_words = 75)
title("Low-rated Reviews")
par(mfrow = c(1, 1))
```

The two clouds tell different things. High-rated reviews are dominated by structural and positive-outcome language: *work*, *help*, *better*, *medic* (medication), *life*, *doctor*, *recommend*, *well*. These are reviews where patients are describing positive experiences and talking about how the drug fits into their life. Low-rated reviews share some vocabulary but add a distinctive negative register: *bad*, *sever* (severe), *horribl* (horrible), *bleed*, *cramp*, *headach*, *wors* (worse), *caus* (cause), *mood*. The word *side_effect* is prominent in both clouds, but more central in low-rated reviews, where negative experiences are disproportionately framed around adverse reactions.
