---
title: "Drug Review Dataset — Phase 0 & Phase 2"
author: "Kaggle University Club Hackathon"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: '3'
subtitle: Data Cleaning & Shared Feature Extraction
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

------------------------------------------------------------------------

# Overview

This notebook covers two phases of our pipeline:

-   **Phase 0** — Load the raw train/test CSVs, combine them, and clean the data
-   **Phase 2** — Engineer a shared set of features that feed into both modelling branches (Aspect 1: patient perspective, Aspect 2: clinical perspective)

The output is a single enriched dataset (`drug_reviews_phase2.csv`) ready for Phase 3 modelling.

------------------------------------------------------------------------

# Setup — Install & Load Packages

```{r packages}
required_packages <- c(
  "tidyverse",   # data wrangling & piping
  "tidytext",    # text tokenisation and TF-IDF
  "sentimentr",  # sentence-level sentiment scoring
  "textclean",   # HTML entity stripping
  "lubridate",   # date parsing
  "quanteda",    # advanced text features
  "ggplot2",     # visualisation
  "scales",      # axis formatting in plots
  "caret"        # createDataPartition for Phase 3
)

installed <- rownames(installed.packages())
for (pkg in required_packages) {
  if (!pkg %in% installed) install.packages(pkg, dependencies = TRUE)
}

library(tidyverse)
library(tidytext)
library(sentimentr)
library(textclean)
library(lubridate)
library(quanteda)
library(ggplot2)
library(scales)
library(caret)
library(lexicon)
library(quanteda.textstats)
```

```{r}
source("TAB_dfm.R")
```

------------------------------------------------------------------------

# Phase 0 — Data Loading & Cleaning

## Step 0.1 — Load & Combine Train and Test

We combine train and test into one full dataset **before** feature engineering. This ensures all features are computed consistently across the entire dataset. We will re-split into Train/Test in Phase 3 before modelling.

```{r load-data}
train_raw <- read.csv("drugsComTrain_raw.csv", stringsAsFactors = FALSE)
test_raw  <- read.csv("drugsComTest_raw.csv",  stringsAsFactors = FALSE)

# Tag origin so we can re-split later if needed
train_raw$split <- "train"
test_raw$split  <- "test"

# Combine
df <- rbind(train_raw, test_raw)

cat("Combined rows:", nrow(df), "\n")
cat("Columns:", paste(names(df), collapse = ", "), "\n")
```

**Quick preview of the raw data:**

```{r preview-raw}
head(df, 5)
```

```{r}
# Exact duplicate review text
dup_review_drug_rows <- df %>%
  add_count(drugName, review, name = "dup_n") %>%
  filter(!is.na(review), dup_n > 1) %>%
  select(uniqueID, drugName, review, usefulCount, dup_n) %>%
  arrange(review, desc(usefulCount))

dup_review_drug_rows
```

## Step 0.2 — Remove Duplicates

```{r remove-duplicates}
n_before <- nrow(df)

df <- df %>%
  distinct(uniqueID, .keep_all = TRUE) %>%       # remove exact duplicate IDs
  distinct(drugName, review, .keep_all = TRUE)    # remove same review on same drug

cat("Rows removed:", n_before - nrow(df), "\n")
cat("Rows remaining:", nrow(df), "\n")
```

```{r}
head(df)
```

## Step 0.3 — Strip HTML from Reviews

Reviews from Drugs.com contain HTML entities (e.g. `&#039;` → `'`) and occasional HTML tags. We clean these out.

```{r strip-html}

df$review <- df$review %>%
  gsub('&#039;', "'",  .) %>%
  gsub('&amp;',  '&',  .) %>%
  gsub('&quot;', '"',  .) %>%
  gsub('&lt;',   '<',  .) %>%
  gsub('&gt;',   '>',  .) %>%
  gsub('&nbsp;', ' ',  .) %>%
  gsub('<[^>]+>', ' ', .) %>%
  str_squish()

# Check a cleaned example
cat("Example cleaned review:\n")
cat(df$review[2], "\n")
```

```{r}

nrow(df)
df

```

## Step 0.4 — Lowercase

```{r lowercase}
df$review_clean <- tolower(df$review)
```

```{r}
head(df)
```

## Step 0.5 — Filter Short Reviews (\< 20 words)

Reviews with fewer than 20 words carry too little signal for NLP analysis.

```{r filter-short}
df$word_count_raw <- str_count(df$review_clean, "\\S+")

n_before <- nrow(df)
df <- df %>% filter(word_count_raw >= 20)

cat("Rows removed (too short):", n_before - nrow(df), "\n")
cat("Rows remaining:", nrow(df), "\n")
```

## Step 0.6 — Handle Missing Conditions

```{r missing-conditions}
cat("Missing conditions before:", sum(is.na(df$condition) | df$condition == ""), "\n")

df$condition[is.na(df$condition) | df$condition == ""] <- "Unknown"

cat("Missing conditions after:", sum(is.na(df$condition) | df$condition == ""), "\n")
```

## Step 0.7 — Parse Date Column

```{r parse-date}
df$date <- dmy(df$date)   # converts "20-May-12" → 2012-05-20

cat("Date range:", format(min(df$date)), "to", format(max(df$date)), "\n")
```

## Step 0.8 — Validate Rating Range (1–10)

```{r validate-rating}
n_before <- nrow(df)
df <- df %>% filter(rating >= 1 & rating <= 10)

cat("Rows removed (invalid rating):", n_before - nrow(df), "\n")
cat("Rating range:", min(df$rating), "to", max(df$rating), "\n")
```

## Phase 0 Summary

```{r phase0-summary}
cat("=== Phase 0 Complete ===\n")
cat("Final cleaned dataset:", nrow(df), "rows x", ncol(df), "columns\n")
```

**Rating distribution after cleaning:**

```{r rating-plot, fig.width=7, fig.height=4}
ggplot(df, aes(x = factor(rating))) +
  geom_bar(fill = "#2c7bb6", colour = "white") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Rating Distribution",
    x     = "Star Rating (1–10)",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

# Phase 1 — EDA

## Review Length

```{r}
df <- df %>%
  mutate(
    review_nchar  = nchar(review_clean),
    review_nwords = str_count(review_clean, "\\S+"),
    rating_group = case_when(
      rating <= 4 ~ "Low",
      rating <= 7 ~ "Medium",
      rating >= 8 ~ "High"
    ) %>% factor(levels = c("Low", "Medium", "High"))
  )
```

```{r}
ggplot(df, aes(x = review_nwords)) +
  geom_histogram(binwidth = 10, fill = "#66c2a5", colour = "white") +
  coord_cartesian(xlim = c(0, 300)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Distribution of Review Length",
    x = "Number of Words",
    y = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
ggplot(df, aes(x = rating_group, y = review_nwords)) +
  geom_boxplot(fill = "#8da0cb", outlier.alpha = 0.12) +
  coord_cartesian(ylim = c(0, 300)) +
  labs(
    title = "Review Length by Rating Group",
    x = "Rating Group",
    y = "Number of Words"
  ) +
  theme_minimal(base_size = 13)
```

## Usefulness Distribution

```{r usefulness-vars}
usefulness_cutoff <- quantile(df$usefulCount, 0.75, na.rm = TRUE)

df <- df %>%
  mutate(
    useful_binary = if_else(usefulCount >= usefulness_cutoff, "Useful", "Less useful"),
    useful_binary = factor(useful_binary, levels = c("Less useful", "Useful"))
  )

cat("Top-25% usefulness cutoff:", usefulness_cutoff, "\n")
```

```{r}
ggplot(df, aes(x = usefulCount)) +
  geom_histogram(binwidth = 5, fill = "#fc8d62", colour = "white") +
  coord_cartesian(xlim = c(0, 100)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Distribution of usefulCount",
    x = "Useful Votes",
    y = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
ggplot(df, aes(x = useful_binary)) +
  geom_bar(fill = "#e6ab02") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Useful vs Less Useful Reviews",
    x = NULL,
    y = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
df <- df %>%
  mutate(
    length_group = cut(
      review_nwords,
      breaks = c(0, 25, 50, 100, 150, 200, Inf),
      labels = c("1-25", "26-50", "51-100", "101-150", "151-200", "200+"),
      include.lowest = TRUE
    )
  )

ggplot(df, aes(x = length_group, y = usefulCount)) +
  geom_boxplot(fill = "skyblue", outlier.alpha = 0.08) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    title = "Useful Votes by Review Length Group",
    x = "Review Length Group",
    y = "usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
ggplot(df, aes(x = rating_group, y = usefulCount)) +
  geom_boxplot(fill = "#fdb462", outlier.alpha = 0.12) +
  coord_cartesian(ylim = c(0, 50)) +
  labs(
    title = "Useful Votes by Rating Group",
    x = "Rating Group",
    y = "usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

## Conditions

```{r top-conditions-table}
top_conditions <- df %>%
  count(condition, sort = TRUE) %>%
  slice_head(n = 15)

top_conditions
```

```{r}
ggplot(top_conditions, aes(x = reorder(condition, n), y = n)) +
  geom_col(fill = "#5e81ac") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 15 Conditions by Number of Reviews",
    x = NULL,
    y = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
condition_usefulness <- df %>%
  filter(condition %in% top_conditions$condition) %>%
  group_by(condition) %>%
  summarise(
    avg_useful = mean(usefulCount, na.rm = TRUE),
    median_useful = median(usefulCount, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

ggplot(condition_usefulness, aes(x = reorder(condition, avg_useful), y = avg_useful)) +
  geom_col(fill = "#d08770") +
  coord_flip() +
  labs(
    title = "Average Useful Votes by Condition",
    x = NULL,
    y = "Average usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

## Sentiment by Rating

```{r}
df <- df %>%
  mutate(
    BLsentiment = review_clean %>%
      get_sentences() %>%
      sentiment_by() %>%
      pull(ave_sentiment)
  )
```

```{r}
ggplot(df, aes(x = factor(rating), y = BLsentiment)) +
  geom_boxplot(fill = "skyblue2", outlier.alpha = 0.12) +
  labs(
    title = "Sentiment Score by Rating",
    x = "Rating",
    y = "Average Sentiment"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
ggplot(df, aes(x = rating_group, y = BLsentiment)) +
  geom_boxplot(fill = "mediumpurple1", outlier.alpha = 0.12) +
  labs(
    title = "Sentiment Score by Rating Group",
    x = "Rating Group",
    y = "Average Sentiment"
  ) +
  theme_minimal(base_size = 13)
```

### Sentiment by Usefulness

```{r}
ggplot(df, aes(x = useful_binary, y = BLsentiment)) +
  geom_boxplot(fill = "peachpuff3", outlier.alpha = 0.12) +
  labs(
    title = "Sentiment Score by Usefulness Group",
    x = "Usefulness Group",
    y = "Average Sentiment"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
df <- df %>%
  mutate(
    sentiment_bin = cut(
      BLsentiment,
      breaks = c(-Inf, -0.5, -0.1, 0.1, 0.5, Inf),
      labels = c("Very negative", "Negative", "Neutral", "Positive", "Very positive")
    )
  )

ggplot(df, aes(x = sentiment_bin, y = usefulCount)) +
  geom_boxplot(fill = "skyblue", outlier.alpha = 0.08) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    title = "Useful Votes by Sentiment Group",
    x = "Sentiment Group",
    y = "usefulCount"
  ) +
  theme_minimal(base_size = 13)

```

Correlation Check

```{r}
cor(df$BLsentiment, df$rating, use = "complete.obs")
cor(df$BLsentiment, df$usefulCount, use = "complete.obs")
```

## Wordcloud

```{r}
library(quanteda.textplots)

dfm_high <- TAB_dfm(
  df %>%
    filter(rating_group == "High") %>%
    pull(review_clean),
  ngrams = 1:2,
  stop.words = TRUE,
  min.prop = 0.01
)

dfm_low <- TAB_dfm(
  df %>%
    filter(rating_group == "Low") %>%
    pull(review_clean),
  ngrams = 1:2,
  stop.words = TRUE,
  min.prop = 0.01
)

par(mfrow = c(1, 2))

textplot_wordcloud(dfm_high, max_words = 75)
title("High-rated Reviews")

textplot_wordcloud(dfm_low, max_words = 75)
title("Low-rated Reviews")

par(mfrow = c(1, 1))
```

## Rating–Tone Mismatch

```{r}
df <- df %>%
  mutate(
    tone_match = case_when(
      rating >= 8 & BLsentiment < 0  ~ "High rating, negative tone",
      rating <= 4 & BLsentiment > 0  ~ "Low rating, positive tone",
      TRUE ~ "Aligned / mixed"
    ),
    tone_match = factor(
      tone_match,
      levels = c("Aligned / mixed",
                 "High rating, negative tone",
                 "Low rating, positive tone")
    )
  )

table(df$tone_match)
```

```{r}
ggplot(df, aes(x = tone_match)) +
  geom_bar(fill = "#c97b63") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Frequency of Rating–Tone Mismatch",
    x = NULL,
    y = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
```

```{r}
ggplot(df, aes(x = tone_match, y = usefulCount)) +
  geom_boxplot(fill = "#8ecae6", outlier.alpha = 0.08) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    title = "Useful Votes by Rating–Tone Mismatch Group",
    x = NULL,
    y = "usefulCount"
  ) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
```

```{r}
mismatch_summary <- df %>%
  group_by(tone_match) %>%
  summarise(
    n = n(),
    avg_useful = mean(usefulCount, na.rm = TRUE),
    median_useful = median(usefulCount, na.rm = TRUE),
    avg_words = mean(review_nwords, na.rm = TRUE),
    .groups = "drop"
  )

mismatch_summary
```

## Useful vs Less Useful Top Words

```{r}
dfm_useful_words <- TAB_dfm(
  df %>%
    filter(useful_binary == "Useful") %>%
    pull(review_clean),
  ngrams = 1,
  stop.words = TRUE,
  min.prop = 0.01
)

dfm_less_words <- TAB_dfm(
  df %>%
    filter(useful_binary == "Less useful") %>%
    pull(review_clean),
  ngrams = 1,
  stop.words = TRUE,
  min.prop = 0.01
)
```

```{r}
top_useful_words <- textstat_frequency(dfm_useful_words, n = 15) %>%
  mutate(group = "Useful")

top_less_words <- textstat_frequency(dfm_less_words, n = 15) %>%
  mutate(group = "Less useful")

top_words_usefulness <- bind_rows(top_useful_words, top_less_words)
top_words_usefulness
```

```{r}
ggplot(top_words_usefulness,
       aes(x = frequency, y = reorder(feature, frequency), fill = group)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~group, scales = "free") +
  labs(
    title = "Top Words in Useful vs Less Useful Reviews",
    x = "Frequency",
    y = NULL
  ) +
  theme_minimal(base_size = 13)
```

## Useful vs Less useful top bigrams

```{r}
dfm_useful_bigrams <- TAB_dfm(
  df %>%
    filter(useful_binary == "Useful") %>%
    pull(review_clean),
  ngrams = 2,
  stop.words = TRUE,
  min.prop = 0.005
)

dfm_less_bigrams <- TAB_dfm(
  df %>%
    filter(useful_binary == "Less useful") %>%
    pull(review_clean),
  ngrams = 2,
  stop.words = TRUE,
  min.prop = 0.005
)

top_useful_bigrams <- textstat_frequency(dfm_useful_bigrams, n = 15) %>%
  mutate(group = "Useful")

top_less_bigrams <- textstat_frequency(dfm_less_bigrams, n = 15) %>%
  mutate(group = "Less useful")

top_bigrams_usefulness <- bind_rows(top_useful_bigrams, top_less_bigrams)
top_bigrams_usefulness
```

```{r}
ggplot(top_bigrams_usefulness,
       aes(x = frequency, y = reorder(feature, frequency), fill = group)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~group, scales = "free") +
  labs(
    title = "Top Bigrams in Useful vs Less Useful Reviews",
    x = "Frequency",
    y = NULL
  ) +
  theme_minimal(base_size = 13)
```

## Sickness mention flag

```{r}
sickness_terms <- c(
  "nausea", "headache", "dizziness", "vomiting", "insomnia",
  "cramps", "bleeding", "anxiety", "depression", "pain",
  "weight gain", "weight loss", "fatigue", "rash", "sleep"
)

sickness_pattern <- str_c(sickness_terms, collapse = "|")

df <- df %>%
  mutate(
    mentions_sickness = str_detect(review_clean, regex(sickness_pattern, ignore_case = TRUE)),
    mentions_sickness = if_else(mentions_sickness, "Mentions sickness", "No sickness mentioned")
  )
```

```{r}
ggplot(df, aes(x = mentions_sickness)) +
  geom_bar(fill = "#90be6d") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Reviews Mentioning Sickness",
    x = NULL,
    y = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
ggplot(df, aes(x = mentions_sickness, y = usefulCount)) +
  geom_boxplot(fill = "#f4a261", outlier.alpha = 0.08) +
  coord_cartesian(ylim = c(0, 100)) +
  labs(
    title = "Useful Votes by Sickness Mention",
    x = NULL,
    y = "usefulCount"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
ggplot(df, aes(x = mentions_sickness, y = rating)) +
  geom_boxplot(fill = "#a8dadc", outlier.alpha = 0.08) +
  labs(
    title = "Ratings by Sickness Mentioned",
    x = NULL,
    y = "Rating"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
sickness_summary <- df %>%
  group_by(mentions_sickness) %>%
  summarise(
    n = n(),
    avg_useful = mean(usefulCount, na.rm = TRUE),
    median_useful = median(usefulCount, na.rm = TRUE),
    avg_rating = mean(rating, na.rm = TRUE),
    .groups = "drop"
  )

sickness_summary
```

### Mismatched Reviews

```{r}
high_neg_examples <- df %>%
  filter(tone_match == "High rating, negative tone") %>%
  select(drugName, condition, rating, usefulCount, BLsentiment, review_clean) %>%
  slice_head(n = 5)

low_pos_examples <- df %>%
  filter(tone_match == "Low rating, positive tone") %>%
  select(drugName, condition, rating, usefulCount, BLsentiment, review_clean) %>%
  slice_head(n = 5)

high_neg_examples
```
