---
title: "Drug Review Dataset — Phase 0 & Phase 2"
author: "TAB Group Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: '3'
subtitle: Data Cleaning & Shared Feature Extraction
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

---

# Overview

This notebook covers two phases of our pipeline:

- **Phase 0** — Load the raw train/test CSVs, combine them, and clean the data
- **Phase 2** — Engineer a shared set of features that feed into both modelling branches (Aspect 1: patient perspective, Aspect 2: clinical perspective)

The output is a single enriched dataset (`drug_reviews_phase2.csv`) ready for Phase 3 modelling.

---

# Setup — Install & Load Packages

```{r packages}
required_packages <- c(
  "tidyverse",   # data wrangling & piping
  "tidytext",    # text tokenisation and TF-IDF
  "sentimentr",  # sentence-level sentiment scoring
  "textclean",   # HTML entity stripping
  "lubridate",   # date parsing
  "quanteda",    # advanced text features
  "ggplot2",     # visualisation
  "scales",      # axis formatting in plots
  "caret"        # createDataPartition for Phase 3
)

installed <- rownames(installed.packages())
for (pkg in required_packages) {
  if (!pkg %in% installed) install.packages(pkg, dependencies = TRUE)
}

library(tidyverse)
library(tidytext)
library(sentimentr)
library(textclean)
library(lubridate)
library(quanteda)
library(ggplot2)
library(scales)
library(caret)
library(lexicon)
```

---

# Phase 0 — Data Loading & Cleaning

## Step 0.1 — Load & Combine Train and Test

We combine train and test into one full dataset **before** feature engineering. This ensures all features are computed consistently across the entire dataset. We will re-split into Train/Test in Phase 3 before modelling.

```{r load-data}
train_raw <- read.csv("drugsComTrain_raw.csv", stringsAsFactors = FALSE)
test_raw  <- read.csv("drugsComTest_raw.csv",  stringsAsFactors = FALSE)

# Tag origin so we can re-split later if needed
train_raw$split <- "train"
test_raw$split  <- "test"

# Combine
df <- rbind(train_raw, test_raw)

cat("Combined rows:", nrow(df), "\n")
cat("Columns:", paste(names(df), collapse = ", "), "\n")
```

**Quick preview of the raw data:**

```{r preview-raw}
head(df, 5)
```



## Step 0.2 — Remove Duplicates

```{r remove-duplicates}
n_before <- nrow(df)

df <- df %>%
  distinct(uniqueID, .keep_all = TRUE) %>%       # remove exact duplicate IDs
  distinct(drugName, review, .keep_all = TRUE)    # remove same review on same drug

cat("Rows removed:", n_before - nrow(df), "\n")
cat("Rows remaining:", nrow(df), "\n")
```

```{r}
head(df)
```



## Step 0.3 — Strip HTML from Reviews

Reviews from Drugs.com contain HTML entities (e.g. `&#039;` → `'`) and occasional HTML tags. We clean these out.

```{r strip-html}

df$review <- df$review %>%
  gsub('&#039;', "'",  .) %>%
  gsub('&amp;',  '&',  .) %>%
  gsub('&quot;', '"',  .) %>%
  gsub('&lt;',   '<',  .) %>%
  gsub('&gt;',   '>',  .) %>%
  gsub('&nbsp;', ' ',  .) %>%
  gsub('<[^>]+>', ' ', .) %>%
  str_squish()

# Check a cleaned example
cat("Example cleaned review:\n")
cat(df$review[2], "\n")
```

```{r}

nrow(df)
df

```


## Step 0.4 — Lowercase

```{r lowercase}
df$review_clean <- tolower(df$review)
```

```{r}
head(df)
```


## Step 0.5 — Filter Short Reviews (\< 20 words)

Reviews with fewer than 20 words carry too little signal for NLP analysis.

```{r filter-short}
df$word_count_raw <- str_count(df$review_clean, "\\S+")

n_before <- nrow(df)
df <- df %>% filter(word_count_raw >= 20)

cat("Rows removed (too short):", n_before - nrow(df), "\n")
cat("Rows remaining:", nrow(df), "\n")
```

## Step 0.6 — Handle Missing Conditions

```{r missing-conditions}
cat("Missing conditions before:", sum(is.na(df$condition) | df$condition == ""), "\n")

df$condition[is.na(df$condition) | df$condition == ""] <- "Unknown"

cat("Missing conditions after:", sum(is.na(df$condition) | df$condition == ""), "\n")
```

## Step 0.7 — Parse Date Column

```{r parse-date}
df$date <- dmy(df$date)   # converts "20-May-12" → 2012-05-20

cat("Date range:", format(min(df$date)), "to", format(max(df$date)), "\n")
```

## Step 0.8 — Validate Rating Range (1–10)

```{r validate-rating}
n_before <- nrow(df)
df <- df %>% filter(rating >= 1 & rating <= 10)

cat("Rows removed (invalid rating):", n_before - nrow(df), "\n")
cat("Rating range:", min(df$rating), "to", max(df$rating), "\n")
```

## Phase 0 Summary

```{r phase0-summary}
cat("=== Phase 0 Complete ===\n")
cat("Final cleaned dataset:", nrow(df), "rows x", ncol(df), "columns\n")
```

**Rating distribution after cleaning:**

```{r rating-plot, fig.width=7, fig.height=4}
ggplot(df, aes(x = factor(rating))) +
  geom_bar(fill = "#2c7bb6", colour = "white") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Rating Distribution",
    x     = "Star Rating (1–10)",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

---

# Phase 2 — Shared Feature Extraction

> **Goal:** Add new feature columns to the cleaned dataset. These features are shared across both modelling branches in Phase 3.

## Feature 1 — Review Length

Word count of each cleaned review. Longer reviews tend to be more detailed and informative.

```{r feature-length}
df$review_length <- str_count(df$review_clean, "\\S+")

summary(df$review_length)
```

```{r length-plot, fig.width=7, fig.height=4}
ggplot(df, aes(x = review_length)) +
  geom_histogram(bins = 60, fill = "#1a9641", colour = "white") +
  scale_x_continuous(limits = c(0, 500)) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Distribution of Review Length",
    x     = "Word Count",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

## Feature 2 — Sentiment Score (AFINN)

| `tidytext` | `get_sentiments("afinn")` | Loads the AFINN lexicon into R |
| `tidytext` | `unnest_tokens()` | Tokenises reviews into individual words |

The AFINN lexicon is a pre-built dictionary of **2,477 English words and phrases** that are commonly associated with sentiment and emotion. It was created by **Finn Årup Nielsen** in 2011 and is one of the most widely used tools for sentiment analysis in NLP.

Each word in the dictionary has been **manually rated** by a human on a scale of −5 to +5 
based on how positive or negative it is:

Each word in a review is matched against the AFINN lexicon (scores from −5 to +5 for each word). We sum all word scores to get an overall sentiment score for the review.

Each word is scored from −5 to +5 by AFINN. These word scores are summed across the entire review, so the final score can range well beyond ±5 depending on how long and how emotional the review is.

```{r feature-sentiment}
afinn <- get_sentiments("afinn")

# Tokenise reviews into individual words
tokens <- df %>%
  select(uniqueID, review_clean) %>%
  unnest_tokens(word, review_clean)

# Match words to AFINN scores and sum per review
sentiment_scores <- tokens %>%
  inner_join(afinn, by = "word") %>%
  group_by(uniqueID) %>%
  summarise(sentiment_score = sum(value), .groups = "drop")

# Merge back (reviews with no AFINN matches get score 0)
df <- df %>%
  left_join(sentiment_scores, by = "uniqueID") %>%
  mutate(sentiment_score = replace_na(sentiment_score, 0))

summary(df$sentiment_score)
```

```{r sentiment-plot, fig.width=7, fig.height=4}
ggplot(df, aes(x = sentiment_score)) +
  geom_histogram(bins = 60, fill = "#d7191c", colour = "white") +
  labs(
    title = "Distribution of Sentiment Scores",
    x     = "Sentiment Score (AFINN)",
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13)
```

### Understanding Sentiment Scores

The AFINN lexicon assigns a score to each **individual word** on a scale of −5 to +5.
However, since we **sum all word scores** across an entire review, the final review-level 
score can range well beyond ±5 depending on the length and emotional intensity of the review.

**At the individual word level (−5 to +5):**

| Score | Meaning | Example Words |
|---|---|---|
| +5 | Extremely positive | "outstanding", "superb" |
| +3 | Moderately positive | "great", "excellent", "happy" |
| +1 | Mildly positive | "good", "okay" |
| 0 | Neutral / not in lexicon | "the", "drug", "tablet" |
| −1 | Mildly negative | "bad", "tired" |
| −3 | Moderately negative | "horrible", "painful", "sick" |
| −5 | Extremely negative | "worst", "suicidal", "hate" |

**At the entire review level (summed score):**

| Score | Meaning | What It Likely Looks Like |
|---|---|---|
| Large positive (e.g. +40) | Overwhelmingly positive | Patient is very satisfied, uses words like "amazing", "helped", "grateful" |
| Small positive (e.g. +5) | Slightly positive | Mixed experience but overall okay |
| Zero (0) | Neutral or balanced | Equal positive and negative language, or very plain language |
| Small negative (e.g. −5) | Slightly negative | Some complaints but not extreme |
| Large negative (e.g. −50) | Overwhelmingly negative | Patient describes bad side effects, pain, and frustration |

In the context of this dataset:

- A **positive score** likely means the patient found the drug helpful and is satisfied
- A **negative score** likely means the patient experienced bad side effects or no improvement
- Scores near **zero** are either neutral in tone or contain equal amounts of positive and 
negative language — which connects directly to our **rating-tone mismatch** feature


## Feature 3 — Emotional Intensity

The absolute value of the sentiment score — measures *how emotional* the language is, regardless of direction. A review that is extremely positive OR extremely negative scores high on emotional intensity.

```{r feature-intensity}
df$emotional_intensity <- abs(df$sentiment_score)

summary(df$emotional_intensity)
```

## Feature 4 — Rating–Tone Mismatch

Flags reviews where the **star rating contradicts the sentiment score**. This is a key insight from the EDA phase and feeds into both modelling branches.

- High rating (≥ 7) but **negative** sentiment → mismatch
- Low rating (≤ 3) but **positive** sentiment → mismatch

```{r feature-mismatch}
df$rating_tone_mismatch <- case_when(
  df$rating >= 7 & df$sentiment_score < 0  ~ 1,
  df$rating <= 3 & df$sentiment_score > 0  ~ 1,
  TRUE                                      ~ 0
)

cat("Total mismatches:", sum(df$rating_tone_mismatch), "\n")
cat("Mismatch rate:", round(mean(df$rating_tone_mismatch) * 100, 2), "%\n")
```

**Why mismatches are common in drug reviews**

Side effects vs. effectiveness trade-off — A patient might write very negatively about side effects ("nausea, dizzy, awful") but still give a high star rating because the drug worked for their condition

Emotional venting — Patients who are suffering tend to use very negative language in their writing even if they are broadly satisfied with the drug

```{r mismatch-plot, fig.width=6, fig.height=4}
df %>%
  count(rating_tone_mismatch) %>%
  mutate(label = ifelse(rating_tone_mismatch == 1, "Mismatch", "Consistent")) %>%
  ggplot(aes(x = label, y = n, fill = label)) +
  geom_col(width = 0.5) +
  scale_fill_manual(values = c("Consistent" = "#1a9641", "Mismatch" = "#d7191c")) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Rating–Tone Mismatch",
    x     = NULL,
    y     = "Number of Reviews"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

**How it is used in Aspect 1 — Patient Perspective**
The question is: "What makes a review useful to other patients?"

A mismatched review is confusing and untrustworthy to readers. If someone writes glowing text but gives 2 stars, other patients do not know what to believe. This likely means:

Mismatched reviews receive fewer useful votes — patients distrust inconsistent reviews 
Consistent reviews are perceived as more credible and persuasive. So rating_tone_mismatch is expected to be a negative predictor of usefulness


**How it is used in Aspect 2 — Clinical Perspective**
The question is: "What can we extract from reviews to support medical understanding?"
A mismatch can actually reveal something clinically interesting:

A patient gives high stars but writes negatively → they may have accepted bad side effects because the drug still worked for their condition
A patient gives low stars but writes positively → they may have liked the drug but stopped due to cost, access, or doctor's advice


## Feature 5 — TF-IDF Specificity

We used the tidytext package in R, specifically the bind_tf_idf() function.

Measures how unique and specific the vocabulary in a review is. A high TF-IDF score means the review uses distinctive words (more informative). A low score means the review uses generic filler language.

We compute TF-IDF per review and take the **mean score** as a single summary value.

```{r feature-tfidf}
tfidf_scores <- tokens %>%
  count(uniqueID, word, name = "n") %>%
  bind_tf_idf(word, uniqueID, n) %>%
  group_by(uniqueID) %>%
  summarise(tfidf_specificity = mean(tf_idf), .groups = "drop")

df <- df %>%
  left_join(tfidf_scores, by = "uniqueID") %>%
  mutate(tfidf_specificity = replace_na(tfidf_specificity, 0))

summary(df$tfidf_specificity)
```

### Interpreting the TF-IDF Specificity Output


The TF-IDF values are small decimals, which is **correct and expected**. This is because TF-IDF is a product of two fractions:

- **TF (Term Frequency)** = number of times a word appears in a review ÷ total words 
  in that review → always between 0 and 1
- **IDF (Inverse Document Frequency)** = log(total reviews ÷ number of reviews containing 
  the word) → a small number when a word is common across many reviews

Since TF-IDF = TF × IDF, the resulting values will always be **small decimals**, 
never large numbers like sentiment scores.

**What the values mean:**

| Value | Meaning |
|---|---|
| Min = 0.004 | Very generic vocabulary — uses very common words like "good", "drug", "feel" |
| Median = 0.034 | Typical review — mix of common and some specific words |
| Max = 0.325 | Very specific vocabulary — uses rare, distinctive words, likely detailed medical descriptions |

The output shows a healthy spread from 0.004 to 0.325, meaning the feature has **good variation** across reviews — which is exactly what is needed for a useful predictor in Phase 3 modelling.


## Feature 6 — Condition Category

There are hundreds of unique conditions in the dataset. We group them into broader clinical categories to reduce dimensionality for the models.

```{r feature-condition-cat}

# Remove more dirty data rows (truncated drug combinations)
df <- df %>%
  filter(!str_detect(condition, regex("^min\\)|^ge \\(|^min / |^acial", ignore_case = TRUE)))

df$condition_category <- case_when(
  str_detect(df$condition, regex("depress|anxiety|bipolar|schizo|adhd|mental|panic|ocd|ptsd|insomnia|sleep|obsessive|restless legs|narcolepsy|post traumatic|posttraumatic|premenstrual dysphoric|sedation|motion sickness|borderline personality|anorexia|binge eating|bulimia|psychosis|mania|hypersomnia|tourette|sexual dysfunction|paranoid|agitated|night terror|nightmare|pseudobulbar",
                                  ignore_case = TRUE)) ~ "Mental Health",
  str_detect(df$condition, regex("pain|migraine|arthritis|fibro|ibromyalgia|headache|back|cramp|muscle spasm|osteoporosis|inflammatory|ankylosing|gout|trigeminal|sciatica|neuralgia|postherpetic|chronic fatigue|benign essential trem|reflex sympathetic|cervical dystonia|atigue|tendonitis|temporomandibular|pruritus|spasticity|tinnitus|dysautonomia",
                                  ignore_case = TRUE)) ~ "Pain & Inflammation",
  str_detect(df$condition, regex("diabetes|thyroid|hormonal|insulin|weight|obesity|metabolic|hypogonad|hashimoto|hyperthyroid|hypothyroid|keratosis|alopecia|edema|hypertriglyceridemia|hyperprolactinemia|dietary supplement|herbal supplement|vitamin|iron deficiency|anemia|hypoestrogenism|precocious puberty",
                                  ignore_case = TRUE)) ~ "Metabolic & Hormonal",
  str_detect(df$condition, regex("birth control|contraceptive|pregnancy|acne|period|menopause|endometri|pcos|uterine|emergency contraception|abnormal uterine|menstrual|postmenopausal|premenstrual|polycystic ovary|ovarian|atrophic vaginitis|hot flash|breast cance|amenorrhea|premature lab|menorrhagia|emale infertility|perimenopausal|melasma|vulvodynia|oophorectomy",
                                  ignore_case = TRUE)) ~ "Women's Health",
  str_detect(df$condition, regex("blood pressure|heart|cholesterol|cardiac|hypertension|stroke|atrial fibrillation|arrhythmia|supraventricular|tachycardia|angina|deep vein thrombosis|pulmonary embolism|acute coronary|raynaud|atrial flutter|mitral valve|left ventricular",
                                  ignore_case = TRUE)) ~ "Cardiovascular",
  str_detect(df$condition, regex("infection|antibiotic|virus|bacterial|fungal|hiv|herpes|uti|hepatitis|sinusitis|influenza|cold sore|onychomycosis|pneumonia|otitis|strep|diarrhea|cold symptom|dental abscess|tonsillitis|head lice|trichomoniasis|malaria|lyme|giardiasis|swine flu|condylomata|pneumococcal|tinea|oral thrush",
                                  ignore_case = TRUE)) ~ "Infectious Disease",
  str_detect(df$condition, regex("asthma|copd|bronchitis|lung|respiratory|allerg|cough|nasal|congestion|rhinitis|sinus|pulmonary|hiccup",
                                  ignore_case = TRUE)) ~ "Respiratory",
  str_detect(df$condition, regex("cancer|tumor|chemo|oncology|lymphoma|leukemia|basal cell|carcinoma|organ transplant|colorectal cance|melanoma|sarcoma",
                                  ignore_case = TRUE)) ~ "Cancer",
  str_detect(df$condition, regex("crohn|colitis|bowel|ibs|gastr|stomach|acid reflux|gerd|constipation|nausea|vomiting|diverticulitis|hemorrhoid|erosive esophagitis|barrett|indigestion|duodenal ulce",
                                  ignore_case = TRUE)) ~ "Gastrointestinal",
  str_detect(df$condition, regex("epilep|seizure|neuro|parkinson|alzheimer|dementia|multiple sclerosis|vertigo|vertig|myasthenia|meniere",
                                  ignore_case = TRUE)) ~ "Neurological",
  str_detect(df$condition, regex("eczema|psoriasis|dermat|skin|rash|rosacea|hyperhidrosis|dry eye|urticaria|keratoconjunctivitis|sjogren|eye redness|burn|glaucoma|lupus|systemic lupus|uveitis|macular|ocular|wrinkle",
                                  ignore_case = TRUE)) ~ "Skin & Eyes",
  str_detect(df$condition, regex("smoking|nicotine|cessation|opiate|opioid|alcohol|substance|dependence|addiction",
                                  ignore_case = TRUE)) ~ "Substance & Addiction",
  str_detect(df$condition, regex("erectile|prostat|testosterone|benign prostatic|anesthesia|dysuria|connective tissue|immunodeficiency|primary immuno",
                                  ignore_case = TRUE)) ~ "Men's Health",
  str_detect(df$condition, regex("bladder|kidney|renal|urinary|overactive|interstitial cystitis",
                                  ignore_case = TRUE)) ~ "Urological",
  df$condition %in% c("Unknown", "Not Listed / Other",
                       "Not Listed / Othe")             ~ "Unknown",
  TRUE                                                  ~ "Other"
)

# Check remaining Other and its percentage
cat("Remaining Other count:", sum(df$condition_category == "Other"), "\n")
cat("Remaining Other percentage:", 
    round(mean(df$condition_category == "Other") * 100, 2), "%\n")



```

```{r condition-plot, fig.width=8, fig.height=5}
df %>%
  count(condition_category) %>%
  mutate(pct = n / sum(n) * 100) %>%
  arrange(desc(pct)) %>%
  ggplot(aes(x = reorder(condition_category, pct), y = pct, fill = condition_category)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(pct, 1), "%")), hjust = -0.1, size = 3.5) +
  coord_flip() +
  labs(
    title = "Reviews by Condition Category (%)",
    x     = NULL,
    y     = "Percentage of Reviews (%)"
  ) +
  theme_minimal(base_size = 13) +
  scale_y_continuous(limits = c(0, max(df %>% count(condition_category) %>% 
                                mutate(pct = n / sum(n) * 100) %>% pull(pct)) + 5))

```



## Feature 7 — Usefulness Binary Label (Target for Aspect 1)

The `usefulCount` column records how many other patients found a review helpful. We binarise it: the **top 25%** of usefulCount = useful (1), the rest = not useful (0).

This is the **target variable** for Aspect 1 (predicting review usefulness).

```{r feature-useful}
useful_threshold <- quantile(df$usefulCount, 0.75)

df$useful_binary <- ifelse(df$usefulCount >= useful_threshold, 1, 0)

cat("Useful threshold (75th percentile):", useful_threshold, "\n")
cat("Useful reviews (1):", sum(df$useful_binary == 1), "\n")
cat("Not useful reviews (0):", sum(df$useful_binary == 0), "\n")
```



```{r}
ggplot(df, aes(x = usefulCount)) +
  geom_histogram(bins = 50, fill = '#2c7bb6', colour = 'white') +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(labels = comma) +
  geom_vline(xintercept = quantile(df$usefulCount, 0.75), 
             colour = 'red', linetype = 'dashed', linewidth = 1) +
  annotate('text', x = quantile(df$usefulCount, 0.75) + 10, 
           y = 50000, label = '75th percentile', colour = 'red') +
  labs(
    title = 'Distribution of Usefulness Votes',
    x     = 'Useful Vote Count',
    y     = 'Number of Reviews'
  ) +
  theme_minimal(base_size = 13)
```
```{r}
# Check balance at different thresholds
thresholds <- c(0.50, 0.75, 0.90)
for (t in thresholds) {
  threshold_val <- quantile(df$usefulCount, t)
  pct_useful    <- mean(df$usefulCount >= threshold_val) * 100
  cat("Threshold:", t*100, "% |",
      "Min usefulCount:", threshold_val, "|",
      "Useful reviews:", round(pct_useful, 1), "%\n")
}
```

The 50th percentile produces an almost equal split but sets the bar too low — 
a review with just 16 votes is not meaningfully more useful than one with 15. The 90th percentile is too restrictive, leaving only 10% of reviews as useful, which risks class imbalance in the model. 

The **75th percentile strikes the best balance** — selective enough to identify genuinely useful reviews, while retaining 
enough positive examples for the model to learn from.
---

# Phase 2 Summary

```{r phase2-summary}
cat("=== Phase 2 Complete ===\n")
cat("Final dataset dimensions:", nrow(df), "rows x", ncol(df), "columns\n\n")

new_features <- data.frame(
  Feature = c("review_length", "sentiment_score", "emotional_intensity",
              "rating_tone_mismatch", "tfidf_specificity",
              "condition_category", "useful_binary"),
  Description = c(
    "Word count of the cleaned review",
    "Sum of AFINN scores — overall positive/negative tone",
    "Absolute sentiment score — how emotionally strong the review is",
    "1 if star rating contradicts sentiment direction",
    "Mean TF-IDF score — how specific and unique the vocabulary is",
    "Grouped medical condition label (10 categories)",
    "1 if in top 25% of usefulCount (target for Aspect 1)"
  ),
  Used_In = c(
    "Aspect 1 + Aspect 2",
    "Aspect 1 + Aspect 2",
    "Aspect 1 + Aspect 2",
    "Aspect 1 + Aspect 2",
    "Aspect 1 + Aspect 2",
    "Aspect 2",
    "Aspect 1 (target)"
  )
)

knitr::kable(new_features, col.names = c("Feature", "Description", "Used In"))
```

**Preview of the enriched dataset:**

```{r preview-enriched}
df %>%
  select(uniqueID, drugName, condition_category, rating,
         review_length, sentiment_score, emotional_intensity,
         rating_tone_mismatch, tfidf_specificity, useful_binary) %>%
  head(10)
```

---

# Save Output

The enriched dataset is saved and ready for **Phase 3 — Modelling**.

> In Phase 3, this dataset will be split 80/20 into train and test sets before any model is trained.

```{r save}
write.csv(df, "drug_reviews_phase2.csv", row.names = FALSE)
cat("Saved: drug_reviews_phase2.csv\n")
cat("Rows:", nrow(df), "| Columns:", ncol(df), "\n")
```

---

*End of Phase 0 & Phase 2*
